---
title: "Tidyverse_Connin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Polling Results: support for gun control

Feature article: https://projects.fivethirtyeight.com/guns-parkland-polling-quiz/

All polls were taken after February 14, 2018, the date of the school shooting in Parkland, Florida

```{r warning = FALSE}

#load our libraries

library(tidyverse)
library(lubridate)
library(magrittr)
library(anytime)
library(ggforce)

# use readr to import/read csv file

guns<-read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/poll-quiz-guns/guns-polls.csv")

glimpse(guns)

# use (lubridate), mutate (dplyr), relocate (dplyr), str_extract (stringr) to calculate the difference between polling start and end dates and put in numeric form. 

guns$Start<-mdy(guns$Start)
guns$End<-mdy(guns$End)
guns%<>%mutate(Diff_Date = difftime(End, Start))%>%
    relocate(Diff_Date, .after = End)%>%
    mutate(Diff_Date = str_extract(Diff_Date, "^."))%>%
    mutate(Diff_Date = as.numeric(as.character(Diff_Date )))

glimpse(guns)

```

Extend possibility  - accomplish the same task without using lubridate. A hint: if you are not using lubridate, you will need to resolve the following error: x character string is not in a standard unambiguous format.

```{r}

# Create boxplot encoding percent support by party for each question and population

guns%<>%mutate%>%pivot_longer(c("Republican Support", "Democratic Support"), names_to = 'Party', values_to = 'Percent_by_Party')%>%rename(Percent_Total_Support = Support)

plot<-guns%>%
        ggplot(aes(x=Party, y=Percent_by_Party))+
        geom_boxplot(alpha=0.7, fill="blue")+
        facet_grid_paginate(Population~Question, ncol=2, page=1, space="free_y")+
        theme_bw()
```

Extend possibility - rework labels for Question and Party categories so that they are legible.


