---
title: "DATA607 - Tidyverse Assignment"
author: "Michael Ippolito"
date: "3/26/2021"
output:
  html_document:
    df_print: paged
    toc: false
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE}

# Init
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
library(tidyverse)
library(knitr)

```

## Joins in Tidyverse {.tabset}

### Overview

These examples will demonstrate the use of Tidyverse inner and outer (left, right, full) joins. The dataset I used for the demonstration is a partial download of French deaths by département (state) for the year 2019 (https://www.kaggle.com/lperez/death-and-population-in-france-19902019).

``` {r datasets}

# Load the départements dataset, converting the département code to int
df_dept <- read.csv("https://raw.githubusercontent.com/mmippolito/cuny/main/data607/tidyverse/departements.csv", sep = ';')
df_dept$departement_code <- as.integer(df_dept$departement_code)
head(df_dept)

# Load the decedents dataset
# For illustrative purposes, I filtered the table to only deaths of people
# aged 103 or older and who died on 1/1/2019. This give us 8 rows:
df_filt <- read.csv("https://raw.githubusercontent.com/mmippolito/cuny/main/data607/tidyverse/deces_2019_filt.csv", sep = ';')
df_filt$departement_deces <- as.integer(df_filt$departement_deces)
head(df_filt)

```

### Inner joins

Inner joins only choose rows where the selected keyed field is equal in each table.

```{r inner}

# Join the filtered decedents table with the départements table:
df_inner <- df_filt %>% 
  inner_join(df_dept, by = c("departement_deces" = "departement_code"))
df_inner

```

If there hadn't been a corresponding departement in the departements table, that row would have been ommitted in the resulting dataset. For example, after deliberately filter out the département de Meurthe-et-Moselle from the departements table; now only 7 rows are shown.

```{r inner2}

# Filter out one of the départements and run the join again:
df_dept_filt <- df_dept %>% filter(departement != "Meurthe-et-Moselle")
df_inner2 <- df_filt %>% 
  inner_join(df_dept_filt, by = c("departement_deces" = "departement_code"))
df_inner2

```

### Left outer joins

Left outer joins choose all rows from the left-hand table and only the data from the right-hand table where the keyed field is equal.

```{r left_outer}

# Left join the filtered decedents table with the filtered départements table.
# Note how all 8 rows of the left table (decedents) are present, but the first département
# is missing (since we filtered it from the départements table).
df_left <- df_filt %>% 
  left_join(df_dept_filt, by = c("departement_deces" = "departement_code"))
df_left

```

### Right outer joins

Right outer joins choose all rows from the right-hand table and only the data from the left-hand table where the keyed field is equal.

```{r right_outer}

# Right join the filtered decedents table with the filtered départements table.
# Now we have all rows from the départements table, but the decendent data is only
# populated in the rows from the 8 rows in the decendents table.
df_right <- df_filt %>% 
  right_join(df_dept_filt, by = c("departement_deces" = "departement_code"))
df_right

```

### Full outer joins

Full outer joins choose all rows from both the left-hand and right-hand tables.

```{r full_outer}

# Full join the filtered decedents table with the filtered départements table.
# Note how all 8 rows of the left table (decedents) are present, but the first département
# is missing (since we filtered it from the départements table).
# Note how even the row with the missing département is displayed.
df_full <- df_filt %>% 
  full_join(df_dept_filt, by = c("departement_deces" = "departement_code"))
df_full

```
